---
title: "Vignette for the multiDA package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette for the multiDA package}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


# Running the multiDA algorithm

Hey, so you are interested in using the multiDA package for your analysis - awesome! Here are a few examples to get your work going. Loading the package using the ```library(multiDA)``` function, we can run a model on the inbuilt ```SRBCT``` data set as follows:

```{r}
library(multiDA)

vy   <- SRBCT$vy
mX   <- SRBCT$mX
res  <- multiDA(mX=mX, vy=vy, penalty="EBIC", equal.var=TRUE, set.options="exhaustive")
```


If we want to predict class labels, we can use the ```predict`` function in order to do so. In this case, we will find the resubstitution error rate for this dataset using the multiDA algorithm. 

```{r}
vals <- predict(res, newdata=mX)$vy.pred          
rser <- sum(vals!=vy)/length(vy)
rser
```


# Exploring the print and plot functions 

We can use the ```print``` command to look at a quick summary of the model fitted.

```{r}
print(res)
```

Further, we can plot the kernel density estimates of features we are interested in. By default, the ```plot``` function plots the top 10 ranked features. If ```ranked=FALSE``, then the user can specify which features to be plotted (specified by column names).

```{r}
plot(res, ranks = 1)
```

An example using specified features
```{r}
plot(res, ranked=FALSE, features = c("V22", "V122"))
```



# multiDA "broom" functions

## multiDA "glimpse" function

a one row data frame, with quick summaries from the algorithm. In the spirit of the "glance" function from the broom package.

```{r}
glimpse_multiDA(res)
```


## multiDA "tidy" function

This returns a tidy data frame, with key results from the trained multiDA object, namely, a data.frame of significant features and their ranks. In the spirit of "tidy" from the broom package.

```{r}
tidy_res <- tidy_multiDA(res)
head(tidy_res)
```


## multiDA "augment" function

Returns a tidy data frame, returning back the original class, matrix of features, augmented with the paritioning of each feature as given by the algorithm. In the spirit of "augment" from the broom package.

```{r}
augment_res <- augment_multiDA(res)
dim(augment_res) #twice as long!
```

