---
title: "Vignette for the multiDA package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette for the multiDA package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


# Running the multiDA algorithm

Hey, so you are interested in using the multiDA package for your analysis - awesome! Here are a few examples to get your work going. Loading the package using the ```library(multiDA)``` function, we can run a model on the inbuilt ```SRBCT``` data set as follows:

```{r}
library(multiDA)

vy   <- SRBCT$vy
mX   <- SRBCT$mX
res  <- multiDA(mX=mX, vy=vy, penalty="EBIC", equal.var=TRUE, set.options="exhaustive")
```


If we want to predict class labels, we can use the ```predict`` function in order to do so. In this case, we will find the resubstitution error rate for this dataset using the multiDA algorithm. 

```{r}
vals <- predict(res, newdata=mX)$vy.pred          
rser <- sum(vals!=vy)/length(vy)
rser
```


# Exploring the print and plot functions 

We can use the ```print``` command to look at a quick summary of the model fitted.

```{r}
print(res)
```

Further, we can plot the kernel density estimates of features we are interested in. By default, the ```plot``` function plots the top 10 ranked features. If ```ranked=FALSE``, then the user can specify which features to be plotted (specified by column names).

```{r}
plot(res, ranks = 1)
```

An example using specified features
```{r}
plot(res, ranked=FALSE, features = c("V22", "V122"))
```



# multiDA "broom" functions
